---
title: "Homework5"
author: "Yuki Joyama"
date: "2023-12-12"
output: 
  pdf_document:
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message=FALSE,
  warning=FALSE)

library(tidyverse)
library(ggplot2)
library(faraway)
library(gtsummary)
library(flextable)

# import data set
df = state.x77 |> 
  as_tibble(rownames = "state") |> 
  mutate(
    state = as_factor(state)
  ) |> 
  janitor::clean_names()
```

# Problem 1
a) The following table shows the descriptive statistics for all variables of interest in 50 States. 
```{r}
df |> 
  select(-state) |> 
  tbl_summary(
    statistic = list(
      all_continuous() ~ "{mean} / {median} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = all_continuous() ~ 1,
    label = list(
      population ~ "Population",
      income ~ "Income per capita",
      illiteracy ~ "Illiteracy (%)",
      life_exp ~ "Life Expectancy (years)",
      murder ~ "Murder rate (per 100,000)",
      hs_grad ~ "High graduates (%)",
      frost ~ "Number of days below freezing",
      area ~ "Land area (mile^2)"
    )
  ) |> 
  as_flex_table()
```

b) 
```{r}
# normality test by shapiro-wilk test
sw_t = tibble(num = 2:9) |> 
  mutate(
    var = colnames(df)[-1],
    sw = map(num, \(i) shapiro.test(df[[i]])),
    sw_results = map(sw, broom::tidy)
  ) |> 
  unnest(sw_results)

sw_t |> 
  select(var, statistic, p.value) |>
  mutate_at(2:3, round, 3) |> 
  mutate(
    p.value = ifelse(p.value < 0.001, "<0.001", p.value)
  ) |> 
  knitr::kable()
```

The results of Shapiro-Wilk test indicates that variable `population`, `illiteracy`, `murder`, `hs_grad`, and `area` is not normally distributed.  
The histogram and Q-Q plots for these variables are as follows:

```{r fig.width=10, fig.height=16}
df_val = df |> 
  select(population, illiteracy, murder, hs_grad, area)
a = colnames(df_val)[1:5]

par(mfrow = c(5, 2), mar = c(4, 4, 2, 1))

for (i in 1:5) {
  sub <- df_val[[i]]
  
  # Create a new plot for each iteration
  hist(sub, main = paste("Hist. of", a[i], sep = " "), xlab = a[i])
  qqnorm(sub, main = paste("Q-Q Plot of", a[i], sep = " "))
  qqline(sub)
}
```

Given the shape of the histograms, I will log-transform `population`, `illiteracy`, and `area`.

```{r}
df_trans = df |> 
  mutate(
    population = log(population),
    illiteracy = log(illiteracy),
    area = log(area)
  )
```

Now, let's check these histograms.

```{r fig.width=10, fig.height=4}
par(mfrow = c(1, 3))
hist(df_trans$population, main = "Hist. of population", xlab = "population")
hist(df_trans$illiteracy, main = "Hist. of illiteracy", xlab = "illiteracy")
hist(df_trans$area, main = "Hist. of area", xlab = "area")
```

c) Automatic procedures 

d) Criterion-based procedures

e) The LASSO method

f) Compare the subsets from c, d, and e

g) Findings











